
```{r}
# Load the relevant libraries - do this every time
# Note this is pulled in from my standard libraries
library(lubridate)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(tidyverse)
library(ggmap)
library(googledrive)
if (!require('tidyverse'))
  install.packages(tidyverse)
if(!require('maps'))
  install.packages(maps)
```

```{r}
# Note: set up an API key see documentation below
register_google(key = "APIKey", write = TRUE)
```

A very last minute, hacked together R script (apologies Python folks) to get plots set up for the SAB :)

# Africa

```{r}
# I believe this is in East Cameroon
myLocation <- c(lon = 14.547250, lat = 5.570445)

# Note we grab the map using MA for Morocco.  Zoom here isn't great, 
myMap <- get_googlemap(center=myLocation,
                 region='MA',
                 source="google", 
                 maptype="terrain",
                 crop=FALSE, 
                 #force=TRUE,
                 zoom = 3,
                 archiving = TRUE,
                 filename = "Africa", 
                 where = "./maps")

p <- ggmap(myMap)

p
```

Retrieve data from Google Sheet; necessary to capture any updated as needed.

```{r}
# Comment/Uncomment to reload the data from Google (posted now to the WP folder)

drive_download(
  "NA Status-Phase2",
  path = "nodeData.csv",
  overwrite = TRUE
)
```

Read in file:

```{r}
nodeData <- read.csv('nodeData.csv',
                     header=TRUE, colClasses = c("character", "character", "factor", "factor", "numeric", "numeric", "character", "character", "character", "character"),
                     na.strings = c("","NA")
                       )
head(nodeData)
```

Okay the table is a bit of a hack but it's easy to edit.  We have the map, now let's plot just the coordinates:

```{r}
p + geom_point(aes(x = Longitude, 
                   y = Latitude,  
                   colour = "black"), 
               data = nodeData, size = 3) + 
  theme(legend.position="bottom")
```

Let's get a few numbers. How many have attempted node accreditation?

```{r}
nodeData$Attempted <- nodeData[,c("Variant.calling", 
                                  "GWAS", 
                                  "Microbiome", 
                                  "RNA.Seq")] %>% is.na %>% `!` %>% rowSums %>% as.integer
sum(nodeData$Attempted > 0)
```

How many have not attempted?

```{r}
sum(nodeData$Attempted == 0)
```

We're now over 50%! This does include nodes expressing interest but who haven't started yet (a few in this group).

How many are pending?

```{r}
apply(X=nodeData[,c("Variant.calling", "GWAS", "Microbiome", "RNA.Seq")],
      MARGIN = 2, FUN = function(y) length(grep(y, pattern = "Pending"))) 
```

No one wants to take GWAS :(

```{r fig.width=6, fig.height=6}
#png("africaAccred1.png", width = 6, height = 6, units = "in")

#nodeData %>% dplyr::filter(Attempted > 0) -> taken
#nodeData %>% dplyr::filter(Attempted == 0) -> notTaken

p2 <- p + 
  # geom_point(aes(x = Longitude, y = Latitude),
  #              data = dplyr::filter(nodeData,Attempted == 0), size = 8, color = "black") + 
  geom_point(aes(x = Longitude, y = Latitude,
                   color = as.factor(Attempted)),
               data = dplyr::filter(nodeData, Attempted > 0), size = 16) + 
  theme(legend.position=c(0.5,0.05),
        legend.direction="horizontal",
        legend.title=element_text(size=16),
        legend.text=element_text(size=16)) + 
  scale_fill_brewer(palette="Dark2") +
  xlab(element_blank()) + ylab(element_blank()) +
  labs(size = element_blank(),
       colour = "Num. exercises")

p2

#dev.off()
```

```{r fig.width=6, fig.height=6}
foo <- dplyr::filter(nodeData,Attempted == 0)
foo <- dplyr::filter(foo, Country != 'USA')
p2 + geom_point(aes(x = Longitude, y = Latitude),
               data = foo, size = 8, color = "black")
```

This handles the border issue with Morocco.

Here is Azza's version, with the dots overlayed only for those who are showing up in the 

```{r fig.width=4, fig.height=4}
mapWorld <- map_data('world') %>%
  mutate(Countries = 'world') %>%
  filter(!str_detect(region, "Antarctica"))
bordersWorld <- borders('world', colour="gray50") # create a layer of borders

consort <- c('Morocco', 'Tunisia','Egypt', 'Sudan','Mali', 'Senegal',
             'Nigeria', 'Ghana', 'Uganda', 'Kenya', 'Tanzania', 'Botswana', 'Malawi', 'Zimbabwe',
             'South Africa', "Mauritius")
mapConsortia <- map_data("world", regions = consort) %>%
  mutate(Countries = 'H3ABioNet Consortium')
```

```{r fig.width=4, fig.height=4}
p <- ggplot() + 
  geom_polygon(data = mapWorld, aes(x=long, y=lat, group=group), fill = 'seashell1',
               color = 'gray50') +
  # geom_polygon(data = mapConsortia, aes(x=long, y=lat, group=group, fill = Countries),
  #              alpha = 1, color = 'gray50') + 
  coord_cartesian(xlim = c(-25, 60), ylim = c(-40, 50)) #+
  # theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(),
  #       legend.position = "bottom", panel.grid = element_blank())  
p
```

```{r fig.width=6, fig.height=6}
p <- p + 
  # geom_point(aes(x = Longitude, y = Latitude),
  #              data = dplyr::filter(nodeData,Attempted == 0), size = 8, color = "black") + 
  geom_point(aes(x = Longitude, y = Latitude,
                   color = as.factor(Attempted)),
               data = dplyr::filter(nodeData, Attempted > 0), size = 16) + 
  theme(legend.position=c(0.5,0.05),
        legend.direction="horizontal",
        legend.title=element_text(size=16),
        legend.text=element_text(size=16)) + 
  #scale_fill_brewer(palette="Dark2") +
  xlab(element_blank()) + ylab(element_blank()) +
  labs(size = element_blank(),
       colour = "Num. exercises")
p
```

```{r fig.width=6, fig.height=6}
p + geom_point(aes(x = Longitude, y = Latitude),
               data = dplyr::filter(nodeData,Attempted == 0), size = 8, color = "black")
```

Now we're playing around.

Riaan's suggestion, `leaflet`:

```{r}
if (!require('leaflet'))
  install.packages('leaflet')
```

```{r}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=14.547250, lat=-5.570445)
m
```

```{r}
m %>% addProviderTiles(providers$Stamen.Toner)
```

## More plots

Let's collapse the table down

```{r}
years <- nodeData[,c("Variant.calling", "GWAS", "Microbiome", "RNA.Seq")] %>% 
  apply(2, function(x) { sub('.*\\s\\((\\d+)\\)', "\\1", x) })

tot.yrs <- data.frame("Year" = 2014:2019)

# TODO: redo as apply?
for (acc in colnames(years)) {
  yr.tbl <- as.data.frame(table(years[,acc]))
  colnames(yr.tbl) <- c("Year", acc)
  
  yr.tbl$Year <- as.integer(as.character(yr.tbl$Year))
  tot.yrs <- left_join(tot.yrs, yr.tbl, by = "Year")
}

tot.yrs[is.na(tot.yrs)] <- 0

#things are being coersced to double, these should be int
tot.yrs[,2:5] <- sapply(tot.yrs[,2:5], as.integer)
tot.yrs
```


```{r}
# gather is not working here?
tmp <- tot.yrs %>% reshape2::melt("Year", variable = "Exercise", value.name = "Total")

ggplot(tmp, aes(x = Year, y = Total)) +  geom_col(aes(fill=Exercise)) +
  scale_x_continuous("Year", breaks = 2014:2019, labels=2014:2019) +
  scale_y_continuous("Total", breaks = 1:11, labels=1:11, minor_breaks = FALSE)

```

